---
title: "FNCE611 Problem Set 1"
author: "Jessica Blum, Jordan Farrer, Marc Lanthemann"
date: '2017-02-01'
output: function(...) {

          fmt <- rmarkdown::pdf_document(toc = TRUE, number_section = TRUE, df_print = 'kable',...)
        
          fmt$knitr$knit_hooks$size = function(before, options, envir) {
            if (before) return(paste0("\n \\", options$size, "\n\n"))
            else return("\n\n \\normalsize \n")
          }
        
          return(fmt)
        }

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', size = 'small')
```

# APR and EAR

$$1 + EAR = \bigg(1 + \frac{APR}{m}\bigg)^m $$

$$EAR = \bigg(1 + \frac{0.15}{12}\bigg)^{12} -1 =16.08\%$$

# Present Value / Future Value

$$PV = \frac{C}{(1+r)^t} = \frac{33000}{(1+0.05)^3}= \$28,506.60$$

# Perpetuities

$$PV = \frac{C}{r} = \frac{875}{0.088} = 9943.18 Rupee$$

# Annuities / Mortages

## (a)

$$PV = \frac{C}{\frac{r}{m}}\bigg(1 - \frac{1}{(1+\frac{r}{m})^{tm}}\bigg) = \frac{PV\frac{r}{m}}{\big(1 - \frac{1}{(1+\frac{r}{m})^{tm}}\big)} = \frac{100,000\frac{0.06}{12}}{\big(1 - \frac{1}{(1+\frac{0.06}{12})^{12\times30}}\big)} = \$599.55$$

## (b) Refiance

Present value of remaining payments:

$$PV = \frac{C}{\frac{r}{m}}\bigg(1 - \frac{1}{(1+\frac{r}{m})^{tm}}\bigg) = \frac{599.55}{\frac{0.06}{12}}\bigg(1 - \frac{1}{(1+\frac{0.06}{12})^{25\times 12}}\bigg) = 93,054.30$$
At at 5% interest rate, the payments on the remaining present value for 25 years is

$$C_{25 years} = \frac{PV\times \frac{r}{m}}{1 - \frac{1}{(1+\frac{r}{m})^{t\times m}}}=\frac{93,054.30\times \frac{0.05}{12}}{1 - \frac{1}{(1+\frac{0.05}{12})^{25\times 12}}} = 543.99$$

and for 30 years is

$$C_{30 years} = \frac{PV\times \frac{r}{m}}{1 - \frac{1}{(1+\frac{r}{m})^{t\times m}}}=\frac{93,054.30\times \frac{0.05}{12}}{1 - \frac{1}{(1+\frac{0.05}{12})^{30\times 12}}} = 499.54$$

So, the NPV of the savings considering the $2,000 refiancing fee is

$$ NPV_{25 years} = -2000 + \frac{C_{before}-C_{after}}{{\frac{r}{m}}}\bigg(1 - \frac{1}{(1+\frac{r}{m})^{t \times m}}\bigg) = -2000 +  \frac{599.55-543.99}{{\frac{0.05}{12}}}\bigg(1 - \frac{1}{(1+\frac{0.05}{12})^{25 \times 12}}\bigg) = 7,504$$

$$ NPV_{30 years} = -2000 + \frac{C_{before}-C_{after}}{{\frac{r}{m}}}\bigg(1 - \frac{1}{(1+\frac{r}{m})^{t \times m}}\bigg) = -2000 +  \frac{599.55-499.54}{{\frac{0.05}{12}}}\bigg(1 - \frac{1}{(1+\frac{0.05}{12})^{30 \times 12}}\bigg) = 16,630$$

As the NPV is positive for either scenario, refiancing is the preferred option. 

# U.S. Treasury Securities

## (a) Asking Price

\begin{align}
\ P_{ask} & = \frac{\frac{Coupon}{2}}{\frac{Yield}{2}}\bigg(1-\frac{1}{(1+\frac{Yield}{2})^{2\times years}}\bigg) + \frac{100}{(1+\frac{yield}{2})^{2\times years}} \\
 & = \frac{\frac{0.0215}{2}}{\frac{0.01945}{2}}\bigg(1-\frac{1}{(1+\frac{0.01945}{2})^{2\times 5}}\bigg) + \frac{100}{(1+\frac{0.01945}{2})^{2\times 10}} = 100.85
\end{align}


## (b) Duration

\begin{align}
\ duration &= \frac{\sum_{i=1}^{N}\frac{t_i\,{cf}_i}{(1+y)^i}}{\sum_{i=1}^{N}\frac{cf_i}{(1+y)^i}} \\
 & = \frac{\frac{1\,\frac{0.02125}{2}}{(1+\frac{0.01945}{2})^1} + \frac{2\,\frac{0.02125}{2}}{(1+\frac{0.01945}{2})^2} + \dotsm + \frac{10\,\frac{1.02125}{2}}{(1+\frac{0.01945}{2})^{10}}}{\frac{\frac{0.02125}{2}}{(1+\frac{0.01945}{2})^1} + \frac{\frac{0.02125}{2}}{(1+\frac{0.01945}{2})^2} + \dotsm + \frac{\frac{1.02125}{2}}{(1+\frac{0.01945}{2})^{10}}} \\
 & = \frac{9.624}{1.009} = 9.54
\end{align}

Half-year duration is $9.54$, so the full year duration is $9.54 \times 0.5 = 4.77$.

# Term Structure

Investing in long term bonds is riskier than investing in short term bonds, therefore interest rates for longer term bonds will increase (at a diminishing rate) and lead to an upward slope

# Corporate Bonds

## Yield to Maturity

If the coupon rate is 5.2%, then every 6 months Whole Foods will pay $1bn \times \frac{0.052}{2} = 26m$. Thus we can calculate the yield to maturity in Dec 2015 as

\begin{equation}
\ PV = \frac{coupon}{(1+\frac{y}{2})^1} + \frac{coupon}{(1+\frac{y}{2})^2} + \dotsm + \frac{FV + coupon}{(1+\frac{y}{2})^{10}}
\end{equation}

\begin{equation}
\ 998.61 = \frac{52}{(1+\frac{y}{2})^1} + \frac{52}{(1+\frac{y}{2})^2} + \dotsm + \frac{1000 + 52}{(1+\frac{y}{2})^{10}}
\end{equation}

where

```{r}
ytm <- function(price, coupon, term, freq, prin = 100) {
    npv <- function(yld, term, cpn, freq, prin) {
        i <- 1:(term * freq)
        payment <- (coupon / freq) * prin
        disc <- (1 + yld / freq)
        sum(payment / disc^(i)) + prin /(disc^(term * freq))
    }
    uniroot(function(yld) npv(yld, term, cpn, freq, prin) - price,
interval=c(0,100), tol=0.000001)$root
}

y = ytm(price = 998.61, coupon = 0.052, term = 5, freq = 2, prin = 1000)
```



$$y = `r round(100*y, 3)`\%$$

## Spread

$$Spread = ytm_{Coupon} - 3\%$$

$$5.23\% - 3\% = 2.23\%$$